\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-02-23 trunk/europa.tex
updates on europa.tex for Javier

@@

Open Conditions and Threats allow flaw detection and resolution at a
higher-level of abstraction (i.e. in terms of objects and tokens) than
that of simply binding variables as is common in Constraint
Satisfaction Problems. This is advantageous when one applies
heuristics for ordering choices since it provides a richer context in
which to make decisions.  Furthermore, it aids in reducing the amount
of work done by a solver so that only the necessary refinements are
made, otherwise leaving the partial plan with maximum flexibility. For
example, one can omit unbound variables which are time-points of
tokens since threats will force a solver to impose restrictions on
these variables based on the semantics of the objects to which their
tokens apply. Thus the planning process may yield partially-ordered
plans for which all possible extensions are provably valid.

The Plan Space Planning algorithm described above can be implemented
in many different ways depending on the approach chosen for flaw and
resolver selection and for backtracking. \eus built-in solver
implements a chronological backtracking algorithm that is summarized
in Algorithm \ref{alg:europa:solve}.

@@

The algorithm takes as input a partial plan p and returns true if a
complete and consistent refinement of p could be found (or if p is
initially complete and consistent), and false otherwise.

\item Line \ref{li:propagate}: Propagate the constraints to test for
  inconsistency. If found to be inconsistent, then we can return false
  since this is a dead-end i.e. no refinements to p can yield a
  consistent plan.

\item Line \ref{li:selectflaw}: Choose a flaw from the set of
  available flaws.

\item Line \ref{li:noflaw}: If there are no flaws, then p is complete
  and we can terminate with success.

\item Line \ref{li:decisionpt}: Otherwise we formulate a decision
  point which is a branch in the search space. Each choice is a
  particular refinement operation and the DecisionPoint collects all
  possible refinement operations for the given flaw.

\item Line \ref{li:decisionloop}: We keep trying until the chosen flaw
  is resolved, or until we run out of resolvers to try.

\item Line \ref{li:decide}: A new partial plan is obtained by
  application of a refinement operator. Note that the ordering over
  refinement operators to select is a non-deterministic step.

\item Line \ref{li:recurse}: Recursive call to solve the new planning
  problem. If successful, then we are done.

\item Line \ref{li:backtrack}: Otherwise, retract the last refinement
  operation and move on to try the next one.

\item Line \ref{li:noplan}: If we arrive here, then we have exhausted
  all options to resolve the flaw, including the case where no options
  were available initially. Thus the problem cannot be solved.

This algorithm provides for a sound and complete search, assuming that
no flaws or available refinement operators are pruned unnecessarily.
Dead-ends in the search are discovered through constraint propagation.
Constraint propagation is a vehicle for evaluating the consistency of
a partial plan and also for filtering infeasible values from
consideration prior to commitment, allowing in some cases a strong
look-ahead capability which is essential for tractable
search. Consistency testing is initiated by the isConsistent
procedure. The algorithm permits a heuristically controlled search by
applying orderings for chooseFlaw and makeDecisionPoint. It results in
a chronologically-backtracking search.

Heuristics can be implemented by applying ordering/filtering policies
for \texttt{chooseFlaw} and \texttt{makeDecisionPoint}, for instance,
by dealing with earliest flaws first, or by making planning decisions
(open condition flaws) before scheduling ones (threats), etc.

It should be emphasized that while this algorithm is commonly
employed, it is only one of many that could be implemented. As a
simple example, consider the following solver for the N-Queens
problem, where there is an integer variable Vi to represent the queen
in column i, so that for instance, V1=3 means that the variable on
column 1 is placed on row 3.  non-attack constraints are posted as
arithmetic constraints among the Vi variables to ensure that all the
rows and diagonals contain only one queen.

@@

\eus Constraint Reasoning Engine can be used to compute the non-attack
constraint violations (\texttt{ce.getViolations()} is a straight call
on the \eu API), to easily pick the queen that has he most violations
by implementing \texttt{getQueenWithMaxViolation(ce)} and to look
ahead and see what moves are possible for a queen that would remove
its violations by implementing
\texttt{getMoves(ce,queenToMove,curPos)}.

>>>>

\subsubsection{Flaw Resolution}
\label{sec:europa:flaws}

@@

Open conditions and threats allow flaw detection and resolution at a
higher-level of abstraction (\ie in terms of objects and tokens) than
that of binding variables as is common in Constraint Satisfaction
Problems. This is advantageous when a solver applies heuristics for
ordering choices since it provides a richer context in which to make
decisions.  Furthermore, it aids in reducing the amount of work done
by a solver so that only the necessary refinements are made leaving a
partial plan with sufficient flexibility. For example, one can bypass
committing unbound timepoint variables since threats will force a
solver to impose ordering restrictions on these variables.  Thus the
planning process may yield partially-ordered plans for which all
possible extensions are provably valid.

The Plan Space Planning algorithm described above can be implemented
in multiple ways depending on the approach chosen for flaw and
resolver selection and for backtracking. \eus built-in solver
implements a chronological backtracking algorithm that is summarized
in Algorithm \ref{alg:europa:solve}.

@@

The algorithm takes as input a partial plan $p$ and returns true if a
complete and consistent refinement of $p$ could be found (or if $p$ is
initially complete and consistent) and false otherwise. Briefly the
algorithm is described as:

\item[Line \textbf{\ref{li:propagate}}]: Propagate the constraints to
  test for inconsistency. If inconsistent, then return false since no
  refinements to $p$ can yield a consistent plan.

\item[Line \textbf{\ref{li:selectflaw}}]: Choose a flaw from the set
  of available flaws.

\item[Line \textbf{\ref{li:noflaw}}]: If there are no flaws, then $p$
  is complete and terminate with success.

\item[Line \textbf{\ref{li:decisionpt}}]: Formulate a decision point
  which is a branch in the search space. Each choice is a particular
  refinement operation and the \texttt{DecisionPoint} collects all
  possible refinement operations for the given flaw.

\item[Line \textbf{\ref{li:decisionloop}}]: Keep trying until the
  chosen flaw is resolved, or until there are no further resolvers to
  try.

\item[Line \textbf{\ref{li:decide}}]: A new partial plan is obtained
  by application of a refinement operator. Note that the ordering over
  refinement operators to select is a non-deterministic step.

\item[Line \textbf{\ref{li:recurse}}]: Recursive call to solve the
  new planning problem. If successful, then success.

\item[Line \textbf{\ref{li:backtrack}}]: Otherwise, retract the last
  refinement operation and move on to try the next one.

\item[Line \textbf{\ref{li:noplan}}]: Arriving here implies all
  options to resolve the flaw have been exhausted, including the case
  where no options were available initially. Thus the problem is
  insoluble.

Algorithm \ref{alg:europa:solve} provides for a sound and complete
search, assuming that no flaws or available refinement operators are
pruned unnecessarily.  Dead-ends in the search are discovered through
constraint propagation, a vehicle for evaluating the consistency of a
partial plan and also for filtering infeasible values from
consideration prior to commitment. % This allows in some cases a strong
% look-ahead capability which is essential for tractable
% search.
Consistency testing is initiated by the \texttt{isInconsistent}
procedure. The algorithm permits a heuristically controlled search by
applying orderings for \texttt{chooseFlaw} and
\texttt{makeDecisionPoint} resulting in a chronologically-backtracked
search. Heuristics can be implemented by applying ordering/filtering
policies for \texttt{chooseFlaw} and \texttt{makeDecisionPoint}, for
instance, by dealing with earliest flaws first, or by making planning
decisions (open condition flaws) before scheduling decision (threats).

It should be emphasized that while this algorithm is commonly employed
with \eu implementations, it is only one of many that could be
implemented. As a simple example, consider the following solver for
the $N$-Queens problem, where an integer variable V$_i$ represents the
queen in column $i$; V$_1$ $= 3$ implies that the variable on column
$1$ is placed on row $3$.  Non-attack constraints are posted as
arithmetic constraints among the V$_i$ variables to ensure that all
the rows and diagonals contain only one queen. \eus Constraint
Reasoning Engine or \texttt{CRE} can then be used to compute the
non-attack constraint violations, with \texttt{ce.getViolations()} a
call on the API, to pick the queen that has he most violations.  By
implementing \texttt{getQueenWithMaxViolation()} and looking ahead to
see what moves are possible for a queen that would remove its
violations with the implementing of \texttt{getMoves()}, this solver
provides a viable planner implementation for a range of possible
initial states.

@@
